### 2020-06-02 appendix/major.tex
Initial Overleaf Import

\section{Properties of majorization}
\label{app:major}

\subsection{Equivalent conditions for majorization}

\begin{theorem}
Given $\bmx, \bmy, \bmd \in \mathbb{R}^n$, such that the components of $\bmd$ are positive, the following statements are equivalent:
 \begin{enumerate}
	\item[(TM1)] $\bmx \prec_{\bmd} \bmy$;
	\item[(TM2)] $\Gamma_{\bmd}({\bmx}) \prec \Gamma_{\bmd}({\bmy})$;
	\item[(TM3)]\label{en:tm3} $\sum\limits_{i=1}^n \abs{x_i - t d_i} \leq \sum\limits_{i=1}^n \abs{y_i - t d_i}$ for all $t \in \mathbb{R}$;
	\item[(TM4)] $\sum\limits_{i=1}^n (x_i - t d_i)^+ \leq \sum\limits_{i=1}^n (y_i - t d_i)^+$ for all $t \in \mathbb{R}$ and $\sum\limits_{i=1}^n x_i = \sum\limits_{i=1}^n y_i$;
	\item[(TM5)] $\forall k, L_{\bmx|\bmd}(k) \leq L_{\bmy|\bmd}(k)$ and $L_{\bmx|\bmd}(k=n) = L_{\bmy|\bmd}(k=n)$.
 \end{enumerate}
\end{theorem}
\begin{proof}
    \begin{enumerate}
        \item[1$\leftrightarrow2$]
        Suppose now there exists a stochastic $S$ such that $\bmx = S\bmy$ with $\bmd = S\bmd$ and let $B = \Gamma_{\bmd} \circ S \circ \Gamma_{\bmd}^{-1}$.
        $B$ is a $D$-dimensional bistochastic matrix, since composition of stochastic matrices is stochastic and $(\Gamma_{\bmd} \circ S \circ \Gamma_{\bmd}^{-1}) (\frac{1}{D}\bm{1}) = (\Gamma_{\bmd} \circ S) (\bm{d}) = \Gamma_{\bmd}(\bm{d}) = \frac{1}{D}\bm{1}$. Then, $B$ maps $\Gamma_{\bmd}({\bmy})$ to $\Gamma_{\bmd}({\bmx})$.
        Conversely, given $B$, let $S = \Gamma_{\bmd}^{-1} \circ B \circ \Gamma_{\bmd}$.
        Similarly, $S$ is the stochastic matrix that preserves $\bmd$ and maps $\bmy$ to $\bmx$.
        \item[$2\leftrightarrow3$]\hspace{-5pt}, $2\leftrightarrow4$, $2\leftrightarrow5$ These three statement are equivalent to \nick{blah} respectively for the embedded vectors $\Gamma_{\bmd}({\bmx}), \Gamma_{\bmd}({\bmy})$.
        This is clear by rewriting
        \begin{align}
            \sum\limits_{i=1}^n \abs{x_i - t d_i} &= \sum\limits_{i=1}^n d_i \abs{\frac{x_i}{d_i} - t} = \sum\limits_{i=1}^D \abs{\Gamma_{\bmd}(\bmx)_i - t}, \\
            \sum\limits_{i=1}^n (x_i - t d_i)^+ &= \sum\limits_{i=1}^D (\Gamma_{\bmd}(\bmx)_i - t)^+, \\
            L_{\bmx|\bmd}(k) &= L_{\Gamma_{\bmd}(\bmx)}(k'), \\
            \text{with}\ k&=1,\dots,n\ \text{and}\ k'=1,\dots,D \nonumber
        \end{align} 
        and similarly for the right hand side.
    \end{enumerate}
\end{proof}

\subsection{Mana properties}
Mana monotonicity can be directly seen due to statement~\ref{en:tm3} in~\cref{thm:dmajor} for $t=0$.
Furthermore, mana is additive due to the multiplicative property~\ref{en:w4} of~\cref{thm:wstate}.