### 2020-06-02 appendix/wigner.tex

\section{Properties of the Wigner distribution}
\label{app:wigner}

Here we present important properties of the Wigner distribution that are used throughout the paper.

\begin{proposition}\label{thm:wstate}
    The Wigner distribution of a state $\rho$ is
    \begin{enumerate}%[label=\enlabel{W}{\arabic*}]
        \item\label{en:w1} Real valued: $\W{\rho} \in \mathbb{R}^{d^2}$;
        \item\label{en:w2} Normalised: $\sum_{\bmz \in \pd} \W[\bmz]{\rho}=1$;
        \item\label{en:w3} Bounded: $\abs{\W[\bmx]{\rho}} \leq \frac{1}{d}$.
        \item\label{en:w4} Additive under mixing:

        $\W[\bmx]{\sum_i p_i \rho_i} = \sum\limits_i p_i \W[\bmx]{\rho_i}$;
        \item\label{en:w5} Multiplicative under tensor products:

        $\W[\bmx_A \oplus \bmx_B]{\rho_A \otimes \rho_B} = \W[\bmx_A]{\rho_A}\W[\bmx_B]{\rho_B}$.
	\end{enumerate}
\end{proposition}

\begin{proposition}
    \label{thm:wchannel}
    The Wigner distribution of a $\cptp$ operation $\E: \cal{B}(\hd[d_A]) \mapsto \cal{B}(\hd[d_B])$ is:
    \begin{enumerate}
        \item\label{en:wo1} Real-valued: $\W[\bmy|\bmx]{\E} \in \mathbb{R}$;
        \item\label{en:wo2} Normalised: $\sum_{\bmz \in \pd[d_B]} \W[\bmz|\bmx]{\E} = 1$ for any $\bmx \in \pd[d_A]$;
        \item\label{en:wo3} Bounded: $\abs{\W[\bmy|\bmx]{\E}} \leq \frac{d_A}{d_B}$;
	    \item\label{en:wo4} \nick{Transitive}: $\W[\bmy]{\E(\rho)} = \sum_{\bmz \in \pd[d_A]} \W[\bmy|\bmz]{\E} \W[\bmz]{\rho}$ for any $\bmy \in \pd[d_B]$.
    \end{enumerate}
\end{proposition}
If $d_A = d_B$, and in particular if operation $\E$ maps a Hilbert space onto itself, then the stochasticity condition $\abs{\W[\bmy|\bmx]{\E}} \leq 1$ is satisfied.