### 2020-06-02 sections/abstract.tex

\begin{abstract}
Magic states are essential for achieving universality in fault-tolerant schemes.
Magic resource theories attempt to quantify magic via monotones and thus describe the manipulation of magic states.
Here we introduce the concept of majorization fragments as a more generalised projection of such theories in discrete odd dimensions which allows for more powerful results than what monotones can provide.
Fragments naturally link the symmetries of a gate sequence with conditions on the convertibility between states.
We demonstrate the power of fragments by providing exact conditions for the convertibility of single-copy qutrit magic states as well as stricter distillation bounds than the established mana monotone in any odd dimension.
\end{abstract}